\newpage
\section*{Contributions}

\begin{sloppypar}
\textbf{Adesh Jothi Prasanna} handled data collection by writing download\_nba\_data.py, which queries the NBA API for 
clutch traditional stats, clutch four factor metrics, and regular season stats. He 
also helped develop rq1\_plots.py, producing the scatter plot of regular season win\% vs clutch win\% and the bar chart of 
clutch - regular win\% difference for all 30 teams. He also wrote the RQ1 results section of the report.He also helped in the presentation creation.

\textbf{Krishna Karthik} led data cleaning and preprocessing by writing nba\_merged.py, which selects relevant columns, 
renames fields with clutch/reg prefixes, and merges the three datasets into a single  CSV. He also created 
rq2\_plots.py, building the Pearson correlation heatmap for Research Question 2, 
and wrote the methods section. He also helped in the presentation creation.

\textbf{Sanchit Kaul} helped develop the remaining RQ2 visualizations by writing rq2\_scatter\_quadrant\_plots.py (the 
clutch eFG\% vs win\% scatter plot and the eFG\% vs TOV\% quadrant plot ) and rq2\_dot\_plot.py 
(the dot plot comparing top 10 and bottom 10 clutch teams). He wrote the RQ2 results, 
introduction and abstract in the report. He also helped reviewing the codebase and helped with writing sections and 
bringing everything together in latex. He also helped in the presentation creation.

\textbf{Devarchith Parashara Batchu} conducted the RQ3 analysis by writing rq3\_plots.py (the slope charts comparing 
regular-season vs clutch FG\% and FT\% for the top 10 teams) and rq3\_3pct\_plot.py (the dumbbell plot for 
regular-season vs clutch 3P\%). He also wrote the summary, limitations, and future directions sections and 
helped in the final formatting and assembly of the report and presentation.He also helped in the presentation creation.
\end{sloppypar}
\newpage
