\section{Introduction}

In professional basketball, it’s often believed that you really see what a team is made of when a game is close at the 
end. Those last few minutes of a tight match require players to do everything exactly right; any errors on defence are 
heavily penalised, and any mistake – a lost ball or a failed free throw – becomes very important. While season-long 
results show how good a team generally is, they include both big wins and close games, and so might not show how a team 
reacts when it’s truly under pressure. 

The NBA officially calls “clutch time” the last five minutes of the fourth period, or any extra time, when the 
difference in score is five points or less.  This gives a clear, measurable period to look at how teams do when 
the game is on the line for all thirty of them. It is still not known whether doing well in clutch situations is 
something a team can do consistently, or is just a matter of luck in a small number of very important plays.

This work - When the Game Is on the Line: A Statistical Analysis of NBA Clutch Performance - is an attempt to 
find out if doing well at the end of a game is only because the team is generally strong, or if it is connected 
to particular numbers in the stats. Clutch moments are really different from normal play: the game gets slower, 
teams begin to deliberately foul, coaches make more and more changes, and the stress of the situation can change 
which shots players take. These things mean that how a team performs in clutch situations may not be the same as 
their overall performance throughout the season.

Using three collections of data from the 2024–25 NBA regular season – clutch stats from the usual box scores, 
clutch efficiency stats using four factors, and regular-season stats from the usual box scores – the study answers 
three questions: (1) are the best regular-season teams also the best in clutch time, (2) what team statistics are most 
strongly connected with doing well in clutch situations, and (3) how does shooting change from the regular season to 
clutch time for the teams which are best at clutch play?
The analysis relies on descriptive statistics, Pearson correlation, and visualizations including
scatter plots, bar charts, heatmaps, slope charts, and dumbbell plots. Because the dataset
comprises only thirty teams from a single season, the findings are exploratory and no causal
claims are made.