\section{Summary}

\begin{sloppypar}
The purpose of this project was to analyze NBA clutch performance throughout the 2024-25 regular season using team-level data from NBA.com. The results show that clutch success is linked to general regular-season team quality but not entirely. In close late-game circumstances several teams performed above or below their season long baseline suggesting that clutch performance is a lot more than just team strength.

The most significant statistical finding was the positive correlation between clutch effective field goal percentage and clutch success. The strongest negative correlation was between turnovers and turnover percentage. According to these results, the most successful clutch teams are the ones that score effectively and also limit their mistakes with the ball. The analysis also indicated that clutch teams that excel under pressure don't simply repeat their regular season shooting profile. The percentage of field goal attempts, three-point percentage and free throw percentage varied considerably in regular-season and clutch situations. This indicates that clutch situations are a distinct performance setting instead of simply an extension of regular game conditions.

\subsection{Limitations}

The analysis was performed at the team level and does not consider specific player influence, lineup combinations and the availability of key players in critical situations. The data used in the project was collected from a single season, and may not be universal across different seasons, rosters, and coaching environments. The clutch definition adheres to the NBA standard of the last five minutes inside 5 points, but an alternative definition could produce different results. Clutch situations have a smaller sample than full season play which may make certain team-level outcomes more influenced by short term variation. The correlational nature of the analysis means that the relationships determined must not be interpreted as direct effects.

\subsection{Future Directions}

The analysis could be extended in several ways in future work. If these patterns are steady over time, a multi-season study might provide us with better insights about a team's ability to perform in the clutch. A player-level extension might investigate whether a few stars or lineups contribute more strongly to clutch outcomes compared to others. Also, with additional defensive variables future work could examine how defensive factors improve the explanation of clutch success in basketball. A more formal statistical modeling method could be utilized to move beyond visual and correlational analysis towards inferential results.
\end{sloppypar}
